Parsing Postfix log files

### 2009-02-16 doc/logparser-architecture.tex
Improved the first section. Rearranged remaining sections. Started introduction.

@@

XXX NEEDS AN INTRODUCTION\@.

XXX MERGE SOME OF THE PAPER'S CONTENT IN HERE\@.

@@

@@

XXX UPDATE SECTION REFERENCES BELOW\@: THEY SHOULD ONLY REFER TO THE
ARCHITECTURE, THE IMPLEMENTATION CAN REFER BACK TO HERE\@.

@@

    \item [Framework]  The framework is the structure that actions and
        rules plug into.  It provides the parsing loop, shared data
        storage, loading and validation of rules, storage of results, and
        other support functions.  The framework is described in detail in
        \sectionref{framework in architecture}.

@@

    \item [Rules]  The rules are responsible for classifying inputs,
        specifying the action to invoke and the regex that matches the
        inputs and extracts data.  Rules provide an easily extensible
        method of associating log lines with actions, and are described in
        detail in \sectionref{rules in architecture}.

@@

It may help to think of the rules and actions as components that plug into
the framework.  The framework tries each rule in turn until is finds one
which matches the log line, then invokes the action specified by the rule.

@@

XXX MERGE THE NEXT TWO PARAGRAPHS\@.  PULL CONTENT FROM THE PAPER\@.

Decoupling the parsing rules from the associated actions and framework
allows new rules to be written and tested without requiring modifications
to the parser source code (significantly lowering the barrier to entry for
new or casual users who need to parse new log lines), and greatly
simplifies framework, actions, and rules.  Decoupling also creates a clear
separation of functionality: rules handle low level details of identifying
log lines and extracting data from a log line; actions handle the higher
level details of following the path a mail takes through Postfix,
assembling the required data before storing it, dealing with complications
arising, etc; the framework provides services to actions and stores data.

Decoupling the actions from the framework simplifies both framework and
actions: the framework provides services to the actions, and does not need
to deal with the complications that arise, or the task of reconstructing a
mail's journey through Postfix; actions benefit from having services
provided by the framework, freeing them to concentrate on the task of
accurately reconstructing each mail's journey through Postfix and dealing
with the complications described in \sectionref{complications}.

Separating the rules from the actions and framework makes it possible to
parse new log lines without modifying the core parsing algorithm.  Adding a
new rule with the action to invoke and a regex to match the log lines is
trivial in comparison to understanding a program's entire parsing
algorithm, identifying the correct location to change and making the
appropriate changes.  Bear in mind that the changes must be made without
adversely affecting existing parsing, particularly as there may be edge
cases that are not immediately obvious ---
\sectionref{yet-more-aborted-delivery-attempts} describes a complication
that occurs only four times in \numberOFlogFILES{} log files tested.
Requiring changes to the parser's code also complicates upgrades, as the
changes must be preserved during the upgrade, and may clash with changes
made by the developer (\sectionref{adding new actions in implementation}
describes safely adding new actions).  \parsername{} allows the user to add
new rules to the database without changing the parsing algorithm, unless
the new log lines to be parsed require functionality not already provided
by the algorithm.  If the new log lines do require new functionality, new
actions can be added to the parser without modifying existing actions or
other parts of the algorithm; only in the rare event that the new actions
require support from other sections of the code will more extensive changes
be required.

There is some similarity between the parser's design and William Wood's
\acronym{ATN}~\cite{atns,nlpip}, a tool used in Computational Linguistics
for creating grammars to parse or generate sentences.  The resemblance
between \acronym{ATN} and the parser is accidental, but it is obvious that
the two different approaches share an underlying separation of concerns;
this appears to be a natural division of responsibility and functionality.
\Tableref{Similarities between ATN and this architecture} shows the
similarities.

@@

>>>>

\label{parser architecture}

@@

@@

To avoid cluttering the explanation of the parser architecture with details
of the difficulties encountered while parsing Postfix log files, the
description of the parser architecture has been separated from the
description of \parsername{}, the implementation developed for this
project.  This chapter presents the architecture that has been designed and
developed for this project, beginning with the overall architecture and
design, followed by the three components of the architecture: Framework,
Actions, and finally Rules.  The discussion in this chapter centers on the
theoretical, XXX EXTEND WHEN THE CHAPTER IS FINISHED\@.

The practical difficulties of writing a parser for Postfix log files are
covered in detail in \sectionref{Postfix Parser Implementation}.

@@

@@

    \item [Framework]  The framework is the structure that actions and
        rules plug into.  It provides the parsing loop, shared data
        storage, loading and validation of rules, storage of results, and
        other support functions.

@@

    \item [Rules]  The rules are responsible for classifying inputs; they
        specify the action to invoke and the regex that matches the inputs
        and extracts data.  Rules provide an easily extensible method of
        associating log lines with actions.

@@

For each input the framework tries each rule in turn until it finds a rule
that matches the input, then invokes the action specified by that rule.

@@

Decoupling the parsing rules from their associated actions allows new rules
to be written and tested without requiring modifications to the parser
source code, significantly lowering the barrier to entry for casual users
who need to parse new inputs, e.g.\ part-time systems administrators
attempting to combat and reduce spam; it also allows companies to develop
user-extensible parsers without divulging their source code.  Decoupling
the framework, actions, and rules simplifies all three and creates a clear
separation of functionality: the framework provides services to the
actions, but does not need to perform any tasks specific to the input being
parsed; actions benefit from having services provided by the framework,
freeing them to concentrate on the task of accurately and correctly
processing the information provided by rules; rules handle the low level
details of classifying inputs and extracting data from those inputs.

Separating the rules from the actions and framework makes it possible to
parse new log lines without modifying the core parsing algorithm.  Adding a
new rule with the action to invoke and a regex to match the log lines is
trivial in comparison to understanding a program's entire parser,
identifying the correct location to change, and making the appropriate
changes.  Bear in mind that changes to the parser must be made without
adversely affecting existing parsing, particularly as there may be edge
cases that are not immediately obvious ---
\sectionref{yet-more-aborted-delivery-attempts} describes a complication
that occurs only four times in \numberOFlogFILES{} log files.  Requiring
changes to the parser's source code also complicates upgrades, as the
changes must be preserved during the upgrade, and may clash with changes
made by the developer.  This architecture allows the user to add new rules
without changing the parser, unless the new log lines require functionality
not already provided by the existing actions.  If the new log lines do
require new functionality, new actions can be added to the parser without
modifying existing actions (\sectionref{adding new actions in
implementation} describes how to safely add new actions); only in the rare
event that the new actions require support from other sections of the code
will more extensive changes be required.

There is some similarity between the parser's design and William Wood's
\acronym{ATN}~\cite{atns}, used in Computational Linguistics for creating
grammars to parse or generate sentences.  The resemblance between the two
(shown in \tableref{Similarities between ATN and this architecture}) is
accidental, but it is obvious that the two different approaches share a
similar division of responsibilities, despite having different semantics.

@@

\section{Conclusion}

XXX TO BE WRITTEN\@.
